My advisor, Adrian Roitberg, has been invaluable in this pursuit.
  I often struggled throughout this work and would run into many roadblocks throughout this journey.
  Fortunately, even short meetings with him would allow me to break through them.
He has taught me many lessons, but the most important has probably been to choose my words carefully because presumably the person I'm talking to is listening, and a poor choice of wording might as well be a lie. I can't assume that the person listening will know what I mean.
I also want to thank my committee members, Peter Hirschfeld, Valeria Kleiman, and Xiaoguang Zhang.

  My wife and family, including the in-laws, have been supportive of my Ph.D. pursuit since the beginning.
  Their kindness and understanding have really helped me maintain the course.
  On our daily runs, my wife and I share our thoughts of the day.
  These runs helped me organize my thoughts and added some structure to this work and the defense.

  The Tretiak group at Los Alamos was kind enough to host me for a couple of months to gain acquaintance with their NEXMD software package.
  During this stay, I asked the maintainers about the software and gained first-hand insight into its use.
  I hope my incessant questions were not too much an annoyance.

  Finally, I want to thank all the Roitberg group members who helped me with this work and made it much less lonely.
  I want to give special thanks to Dr. Pancham La Gupta and Dr. Justin Smith, who notoriously helped me with my qualifier while being snowed-in at a bunker outside of Yosemite.
  I want to thank Kavindri Ranasinghe and Christian Deveroux for their company in room 402. 
  I also want to thank Delaram Ghoreishi and Farhad Ramezanghorbani for allowing me to quiz them about the dissertation process.
  Everyone from the group has been exceptional, and I'll forever value the friendships I made there.
